---
title: "Class 05: Data Visulaization with GGPLOT"
author: Jada Ruiz
format: gfm
---
# Our first ggplot

To use the ggplot2 package I first need to have it installed on my computer. 

To install any package we use the 'install.packages()' command. 

Now can we use it? NO! We need first call 'library(ggplot2)'.

```{r}
library(ggplot2)
ggplot()
```

```{r}
mpg
```

Our first plot of displ vs hwy
All ggplot() graphs are made in the same way:

- data + aesthetics + geoms

```{r}
ggplot(mpg) + aes(x=displ, y=hwy) + geom_point()
```

```{r}
ggplot(mpg) + aes(x=displ, y=hwy) + geom_point() + geom_smooth(method= lm, se=FALSE) 
```
# Plot of gene expression data

First read the data from online. 

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

> Q. How many genes are in this dataset?
5196 genes 

What are the colnames?
```{r}
colnames(genes)
```

A first version plot of this data Condition1 vs Condition2.
```{r}
p <- ggplot(genes) + aes(x=Condition1, y=Condition2, col=State) + geom_point() 
p
```

> Q. How many genes are upregulated and down regulated?

```{r}
head(genes)
```

To get at just the State column:
```{r}
table(genes$State)
```

```{r}
table( c("bimm143", "help", "me", "bimm143"))
```
You can save a ggplot underneath a variable and then recall it and add more to the plot that way without a big long chunk of code.

```{r}
p + scale_colour_manual( values=c("blue","gray","red") ) + labs(title = "My nice title", x="No drug", y="Drug")
```




