---
title: "Class 6 R Functions"
author: "Jada Ruiz"
format: gfm
---

Example Student Grades:
```{r}
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)

```

>Q1. Write a function grade() to determine an overall grade from a vector of student homework
assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an NA value) this can be used as a score to be potentially dropped.

Start by using the mean.
```{r}
mean(student1)
```
Find the minimum value of the student's grade because that is what we plan to drop.
```{r}
min(student1)
```
Different type of minimum function tell you what position the minimum value is in.
```{r}
which.min(student1)
student1
```

Get a vector without the minimum value either by hard coding, which wont work for all students, or by using another method the minus index trick.
```{r}
student1[1:7]
student1[-8]
```

So I will combine the output of 'which_min()' with the minus index trick to get the student's score without the lowest value.
```{r}
mean(student1[-which.min(student1)])
```

This will not work for student 2 specifically because student 2 has NA values which student 1 doesn't have so we didn't consider that while writing the code for student 1. 
```{r}
mean(student2[-which.min(student2)])
```

There is a 'na.rm=FALSE' which is by default, putting it equal to 'TRUE' will allow the code to actually work. 
```{r}
mean(student2[-which.min(student2)], na.rm=TRUE)
student2
```
We see this doesn't exactly work for student 3 though because they have a lot of missing grades, more than one NA. 
We must replace all of the NA (missing values) values with zero.
```{r}
student3[is.na(student3)] <- 0
mean(student3[-which.min(student3)])

```

```{r}
student2[is.na(student2)] <- 0
mean(student2[-which.min(student2)])
```
Now turn into a function:
```{r}
grade <- function(x) {
  x[is.na(x)] <- 0
  mean(x[-which.min(x)])
}

```

>Q2. Using your grade() function and the supplied gradebook, Who is the top scoring student overall in the gradebook? 

```{r}
url <- "https://tinyurl.com/gradeinput"
gradebook <- read.csv(url, row.names=1)
```

Heres the gradebook:
```{r}
head(gradebook)
```

Now time to use the 'apply()' function.
```{r}
results <- apply(gradebook, 1, grade)
results
```

Now time to find the highest scoring student:
```{r}
which.max(apply(gradebook, 1, grade))
max(apply(gradebook, 1, grade))
```

>Q3. From your analysis of the gradebook, which homework was toughest on students (i.e. obtained the lowest scores overall? 

```{r}
which.min(apply(gradebook, 2, sum, na.rm=TRUE))
```

>Q4. Optional Extension: From your analysis of the gradebook, which homework was most
predictive of overall score (i.e. highest correlation with average grade score)?

```{r}
mask <- gradebook
mask[is.na(mask)] <- 0

cor(mask$hw5, results)
```
Or use apply:
```{r}
apply(mask, 2, cor, y=results)
```



