---
title: "Class 10 Halloween Candy Mini-Project"
author: "Jada Ruiz"
format: pdf
---

## Importing Candy Data 

```{r}
candy_file <- "candy-data.csv"

candy = read.csv(candy_file, row.names=1)
head(candy)
```

>Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

There are 85 different types of candy.

>Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

There are 38 fruity candy types.

## What is your favorite candy?

```{r}
candy["Twix", ]$winpercent
```

>Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
candy["Snickers", ]$winpercent
```

Snickers is my favorite candy and its win percent is 81.6%

>Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```

The win percent value for Kit Kat is 76.8%

>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

The win percent value for Tootsie Roll Snack Bar is 49.6%

```{r}
library("skimr")
skim(candy)
```

>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

It looks like the `winpercent` column is on a different scale to the majority of the other columns because the mean of that column is quite larger than 0 while all the other columns are less than 0 for their means.

>Q7. What do you think a zero and one represent for the candy$chocolate column?

The zero and one in the `candy$chocolate` represents whether or not the candy contains chocolate. A zero means there is no chocolate, and a one means there is chocolate.

>Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

```{r}
library(ggplot2)
ggplot(candy)+ aes(winpercent) + geom_histogram(bins=10, col="red", fill="orange")
```

>Q9. Is the distribution of winpercent values symmetrical?

The distribution of winpercent values is not symmetrical. 

>Q10. Is the center of the distribution above or below 50%?

The center of distribution is below 50%

>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
chocolate.inds <- as.logical(candy$chocolate)
chocolate.wins <- candy[chocolate.inds,]$winpercent
mean(chocolate.wins)
fruity.inds <- as.logical(candy$fruity)
fruity.wins <- candy[fruity.inds,]$winpercent
mean(fruity.wins)
```

On average, the chocolate candy is ranked higher than fruity candy.

>Q12. Is this difference statistically significant?

```{r}
t.test(chocolate.wins, fruity.wins)
```

Based on the t-test, we can see that this difference is statistically significant. 

## Overall Candy Ranking 

First we will set up code so the candy is represented by a color. 

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
my_cols
```

>Q13. What are the five least liked candy types in this set?

```{r}
head(candy[order(candy$winpercent),], n=5)
```

The five least liked candy types is Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, and Jawbusters.

>Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
head(candy[order(candy$winpercent, decreasing=TRUE),], n=5)
```

The top five candy types are Reeses Peanut Butter cup, Reeses Miniatures, Twix, Kit Kat, and Snickers.

>Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
ggplot(candy) + aes(winpercent, rownames(candy)) + geom_col(fill=my_cols)
ggsave("tmp.png")
```


>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) + aes(winpercent, reorder(rownames(candy), winpercent)) + geom_col(fill=my_cols)
ggsave("tmp.png")
```
Using the plot, answer these questions:

>Q17. What is the worst ranked chocolate candy?

The worst ranked chocolate candy is Sixlets.

>Q18. What is the best ranked fruity candy?

Starburst is the best ranked fruity candy.

## Taking a look at pricepercent

```{r}
library(ggrepel)
# How about a plot of price vs win
ggplot(candy) + aes(winpercent, pricepercent, label=rownames(candy)) + geom_point(col=my_cols) + geom_text_repel(col=my_cols, size=3.3, max.overlaps = 10)
```

>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

The candy with the highest ranked win percent and that is the least amount of money is Reeses Miniatures.

>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

The top 5 most expensive candy types is Nik L Nip, Nestle Smarties, Ring Pop, Hersheys Krackel, and Hersheys Milk Chocolate. The least popular of these is Nik L Nip.

## Exploring the correlation structure

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```

>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Two variables that are anti-correlated are chocolate and fruity.

>Q23. Similarly, what two variables are most positively correlated?

The two variables that are most positively correlated are chocolate and winpercent.

## Principal Component Analysis

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

Now lets plot the PCA

```{r}
plot(pca$x[,1:2])
```

Now add some color:

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
p <- ggplot(my_data) + aes(x=PC1, y=PC2, size=winpercent/100, text=rownames(my_data), label=rownames(my_data)) + geom_point(col=my_cols)
p
```

```{r}
p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

We see that fruity, pluribus, and hard are picked up strongly in PC1 in the positive direction. This makes sense because when we look at the graph of the PC1 vs. PC2 we see that fruity is very prominent in the right side of the graph which is the PC1 side of the graph. 
